Sampling requirements to detect ecosystem change

### 2019-11-06 R/temporal-sampling.Rmd
added intro section and sketch of theoretical results section

---
title: "Effect of temporal sampling"
author: "Rosalie Bruel"
date: "03/10/2019"
output: 
  html_document:
    df_print: paged
    fig_caption: true
    toc: yes
    toc_depth: 3
    toc_float: true
    number_sections: true
editor_options: 
  chunk_output_type: console
  df_print: paged
fontsize: 11pt
---

@@

getpath4data <- function() {
  if(Sys.getenv("USER")=="Rosalie") return("/Volumes/-/Script R/Data-off-GitHub/temporal-sampling/")
  if(Sys.getenv("USER")=="put here your USER") return("replace here with path to data")
  if(Sys.getenv("USER")!="Rosalie"|Sys.getenv("USER")!="put here your USER") stop("You need to get the data and define their location.")
}

@@

```{r echo=FALSE, message=FALSE, warning=FALSE, cache=T}
output <- ut(y=exp(ar1), sample_size = c(20:50), threshold_error = c(2:10), each = 100, type="mean")
names(output)
head(output$convergence_cptmean)
ggplot(output$convergence_cptmean, aes(sample_size,number_sample_added)) + geom_point(alpha=.3)

@@

>>>>

---
title: "Sampling requirements to detect ecosystem change"
author: "Rosalie Bruel and Easton White"
date: "03/10/2019"
output:
  html_document:
    df_print: paged
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
fontsize: 11pt
editor_options:
  chunk_output_type: console
  df_print: paged
---

# Things to do:

- Easton: build theoretical results figure
- Sketch out figures for case studies
- Conceptual diagram 
- Begin writing - probably a shorter piece
- Possible journals: Methods Ecology and Evolution, Biological Conservation, Journal of Environmental Management, Environmental Monitoring and Assessment, BioScience,

# Introduction and setup

- Set up ideas about monitoring and cost of monitoring
- Idea of detecting change over time
- Historical data and ability to choose how much to sample
- Here, we develop a set of tools to ... We first investigate these tools using a simulated-based apporach. We then apply these tools to the detection of change points for a set of lakes...

# Methods and results

@@

getpath4data <- function() {
  if(Sys.getenv("USER")=="Rosalie") return("/Volumes/-/Script R/Data-off-GitHub/temporal-sampling/")
  if(Sys.getenv("USER")=="eastonwhite") return("~/Desktop/Research/soil-temporal-sampling/")
  if(Sys.getenv("USER")!="Rosalie"|Sys.getenv("USER")!="put here your USER") stop("You need to get the data and define their location.")
}

@@

```{r echo=FALSE, message=FALSE, warning=FALSE, cache=T}
output <- ut(y=exp(ar1), sample_size = c(20:50), threshold_error = 5, each = 100, type="mean")
names(output)
head(output$convergence_cptmean)
ggplot(output$convergence_cptmean, aes(sample_size,number_sample_added)) + geom_point(alpha=.3)

@@

# Theoretical results

We began with a theoretical exploration of the determiners

We modeled a simple AR-1 process with a response variable that represents either population size, biodiversity, or some metric of community composition. The model includes temporal variability ($\sigma$), autocorrelation ($\phi$), the size of the shift ($\delta$), and a specific time when the shift occured ($\tau$):

$$
Model here
$$

We then explored how each of these model components affected our ability to detect a change point. We simulate an entire time series, but we also examine the effect of not sampling every year. We specifically study how the amount of samples included and the type of sampling affects the detection probability. In particular, we examine three types of sampling: random, regular, and iterative.
